---
title: "R Notebook"
output: html_notebook
---
Required Libraries
```{r}
library(glmnet)
library(cvms)
library(tidyverse)
library(caret)
library(e1071)
```




Splitting the data into training and testing data frames:
```{r}
set.seed(2021)
whole <- read.csv("life_expectancy_data.csv")

#removing all rows which contain one or more na value
whole = na.omit(whole)

index <- sample(1:nrow(whole), 1319)
who_train <- whole[index, ]
who_test <- whole[-index, ]
```
Removing all rows which contain one or more na value from the training dataset
```{r}
# who_train = na.omit(who_train)
# who_test = na.omit(who_test)
```



** Creating the naive classification rule **
```{r}
X <- split(who_train, who_train$Status)
```

```{r}
# Mean life expectancy of "Developed" countries
ded = mean(X[1]$Developed[["Life.expectancy"]])
# Mean life expectancy of "Developing" countries
ding = mean(X[2]$Developing[["Life.expectancy"]], na.rm=TRUE)
```
Mean of two life expectancy values, which will be our classification rule.
```{r}
rule = (ded + ding)/2
```

Using the naive classifier to predict "Developed" or "Developing" using the testing data:
```{r}
who_test$prediction <- ifelse(who_test$Life.expectancy <rule, "Developing", "Developed")
naiv_corr_pred = subset(who_test, select = c("Status", "prediction"))
```
Evaluating the naive classifier:
```{r}
eval = evaluate(naiv_corr_pred,
                      target_col = "Status",
                      prediction_col = "prediction",
                type = "binomial")
eval
```
Plotting a confusion matrix for the naive classifier:
```{r}
conf_mat = eval$'Confusion Matrix'[[1]]
conf_mat
plot_confusion_matrix(conf_mat)
```
Misclassification Rate of the naive model:
```{r}
mis_naive = 1-eval$Accuracy
mis_naive
```




*** Linear Model ***
Converting data frames into appropriate matrices

Separating dependent and independent variables into distinct data frames
```{r}
# dependent variable (Y)
y_var = subset(who_train, select = ("Status"))
y_var$Status_num <- ifelse(y_var$Status == "Developed", 1, 0)
to_drop = c("Status")
y_var = data.frame(y_var[, !(names(y_var) %in% to_drop)])

# independent variables (X's)
to_drop = c("Status", "Country")
x_vars = who_train[, !(names(who_train) %in% to_drop)]
```

Converting these data frames into matrices to be ingested by the lm function.
```{r}
y_var2 = apply(as.matrix.noquote(y_var), 2, as.numeric)
x_vars2 = apply(as.matrix.noquote(x_vars), 2, as.numeric)
#test_matrix_clean = apply(as.matrix.noquote(hp_king_test_clean), 2, as.numeric)
```

** Logistic Regression **
```{r}
fit <- glmnet(x_vars2, y_var2, family = "binomial")
plot(fit, xvar = "lambda")
```

```{r}
fit.cv <- cv.glmnet(x_vars2, y_var2, family = "binomial")
plot(fit.cv)
```

```{r}
summary(fit.cv, s = c("lambda.min"))
```

```{r}
glm.fit <- glm(y_var2 ~ x_vars2, family = "binomial")
summary(glm.fit)
```


```{r}
# x_vars2 = apply(as.matrix.noquote(who_test), 2, as.numeric)
# x_vars2 = who_test
# probs = predict(glm.fit, x_vars2, type="response")
```


```{r}
#index2 = head(who_test, 300)
index2 = who_test
```

```{r}
y.hat6 = predict(glm.fit, newdata = index2, type="response")
#y.hat6_df = as.data.frame(y.hat6)
#y.hat6_df
```

```{r}
preds <- as.data.frame(ifelse(y.hat6 > 0.5, "Developed", "Developing")) %>% setNames(c("prediction"))
```


now combining the true values of "Status" with the predictions in order to calculate the model's misclassification rate.
```{r}
preds2 = head(preds, 330)
# preds2 = as.data.frame(preds[569:1137,])
who_test_results = cbind(index2$Status, preds2) %>% setNames(c("True", "Prediction"))
```



Evaluating the Logistic classifier:
```{r}
eval = evaluate(who_test_results,
                      target_col = "True",
                      prediction_col = "Prediction",
                type = "binomial")
eval
```
Plotting a confusion matrix for the Logistic Classifer:
```{r}
conf_mat = eval$'Confusion Matrix'[[1]]
conf_mat
plot_confusion_matrix(conf_mat)
```
Misclassification Rate of the Logistic model:
```{r}
mis_log = 1-eval$Accuracy
mis_log
```









